---
title: "PCA_data"
author: "Yuxuan"
date: "15/06/2021"
output: html_document
---

### This is mainly for data wrangling
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```


## 0) Useful Libraries

```{r message=FALSE, warning=FALSE, include=TRUE}


# wrangling/convenience
library(tidyverse)
library(glue)
library(stringr)
library(sf)
library(data.table)
library(hablar)
library(tibble)
# visualization
library(leaflet)
library(mapview); mapviewOptions(platform = 'leafgl')
#library(ggplot2)
#library(RColorBrewer)
#library(scales)
#library(lattice)

# For pretty knitting
library(lemon)
knit_print.data.frame <- lemon_print
knit_print.tbl <- lemon_print
knit_print.summary <- lemon_print


# For PCA
library(missMDA)
library("factoextra")
```
## import data

```{R}
# import amenities (Cultural/Art facilities)
scores_long<- fread(file.path('clean', '/scores_long.csv'))
vancouver_Prox<- fread(file.path('clean', '/vancouver_Prox.csv'))
destinations <- fread(file.path("clean", "vancouver_facilities_2.csv"))
vancouver_db<-fread(file.path('clean', '/vancouver_db.csv'))
```


### Import the dissemination block shape file
```{r}
canada_shape <- st_read("census2016_DBS_shp/DB_Van_CMA/DB_Van_CMA.shp", stringsAsFactors = FALSE)
# select a greater metropolitan area
metropolitan_area <- "Vancouver"

# filter columns and rows
vancouver_shape <- data.frame(canada_shape[which(canada_shape$CMANAME == metropolitan_area), c(1,11,12,13, 28)])


# id to factor
vancouver_shape$DBUID <- as.factor(vancouver_shape$DBUID)
vancouver_shape$CSDNAME <- as.factor(vancouver_shape$CSDNAME)
paste('Rows = ', nrow(vancouver_shape))
head(vancouver_shape)


# 
vancouver_db$id<-as.factor(vancouver_db$id)
left_join(vancouver_db,vancouver_shape,by=c("id"="DBUID"))->df_van
df_van$pop<-as.numeric(df_van$pop)
df_van%>%group_by(CSDNAME)%>%summarise(pop_total=sum(pop,na.rm = T))->pop_csd

```


# number of amenity by area
```{R}
destinations$city_id<-as.factor(destinations$city_id)
vancouver_shape$CSDUID<-as.factor(vancouver_shape$CSDUID)
right_join(vancouver_shape,destinations,by=c("CSDUID"="city_id"))
```


### Join with vancouver_shape
```{R}

vancouver_shape%>%select(DBUID,CSDUID,CSDNAME,CSDTYPE)->CSDNAME
CSDNAME%>%mutate(NAME=paste(CSDNAME,CSDTYPE))->CSDNAME
scores_long$fromId<-as.factor(scores_long$fromId)
CSDNAME$DBUID<-as.factor(CSDNAME$DBUID)

score_csd<-left_join(scores_long, CSDNAME, by = c('fromId' = 'DBUID'))

CSDNAME%>%group_by(NAME,CSDUID)%>%summarise(n=n())->df_n
amn_csd<-left_join(df_n,destinations,by=c("CSDUID"="city_id"))
amn_csd%>%group_by(NAME)%>%count()->amn_csd
colnames(amn_csd)[2]<-"Amenity_num"
paste(CSDNAME$CSDNAME,CSDNAME$CSDTYPE)%>%unique()
```



### Import rental price file with cenus subdivision id and data cleaning

```{R}
van_rental <- read_csv("van_rental.csv")
van_rental%>%select(X1,`Average Rent ($)`,`Vacancy Rate (%)`)->van_rental
colnames(van_rental)<-c("Name","Average_Rent","Vacancy_Rate")
van_rental$Name<-gsub("[()]","",as.character(van_rental$Name))
van_rental$Average_Rent<-gsub("[!*,]", "", van_rental$Average_Rent)
van_rental$Vacancy_Rate<-gsub("[!*]", "", van_rental$Vacancy_Rate)
van_rental%>%convert(num((Average_Rent)))%>%distinct()->van_rental

unique(van_rental$Name)
unique(score_csd$CSDNAME)
```

### compute average csd score by area
```{R}
score_csd%>%group_by(NAME,CSDUID)%>%summarise(csd_score=mean(score,na.rm=T))->score_csd_agg
```


### Compute population area
```{R}
# 
vancouver_db$id<-as.factor(vancouver_db$id)
right_join(vancouver_db,CSDNAME,by=c("id"="DBUID"))->df_van
df_van$pop<-as.numeric(df_van$pop)
df_van%>%group_by(NAME)%>%summarise(pop_total=sum(pop,na.rm = T))->pop_csd
colnames(pop_csd)[1]<-"NAME"
colnames(pop_csd)[2]<-"population"
```


### merge van_rental and score file

```{R}
left_join(score_csd_agg,van_rental,by = c('NAME' = 'Name'))->joined_df
```

## data cleaning
```{R}
# convert to numeric column
columns <-colnames(vancouver_Prox)[-1]
vancouver_Prox[, columns] <- lapply(columns, function(x) as.numeric(vancouver_Prox[[x]]))
```


```{r}
vancouver_shape%>%select(DBUID,CSDUID,CSDNAME,CSDTYPE)->CSDNAME
CSDNAME%>%mutate(NAME=paste(CSDNAME,CSDTYPE))->CSDNAME
scores_long$fromId<-as.factor(scores_long$fromId)
CSDNAME$DBUID<-as.factor(CSDNAME$DBUID)
vancouver_Prox$id<-as.factor(vancouver_Prox$id)

```

```{r}
# joined the data
prox_df<-left_join(vancouver_Prox, CSDNAME, by = c('id' = 'DBUID'))
```




#### group by area, proxy data
```{R}
prox_df%>%group_by(NAME) %>%summarise_if(is.numeric, funs(mean(., na.rm = TRUE)))->df_proxy_area
```

### prepare for merging 
```{R}
joined_df$NAME<-as.factor(joined_df$NAME)
df_proxy_area$NAME<-as.factor(df_proxy_area$NAME)
amn_csd$NAME<-as.factor(amn_csd$NAME)
pop_csd$NAME<-as.factor(pop_csd$NAME)
left_join(joined_df,df_proxy_area, by = c('NAME'='NAME'))->df
left_join(df,pop_csd, by = c('NAME'='NAME'))->df
left_join(df,amn_csd, by = c('NAME'='NAME'))->df_mereged
```


```{R}
## Export checkpoint
write.csv(df_mereged, 'clean/pca_data.csv', row.names = FALSE)

```

