---
title: "Amenity Density Computation"
author: "Luka Vukovic"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

As we did before, we will run r5r to compute how many amenities there are within
a 1km walking distance (ie. network manhattan distance) from each origin
starting point. The number of amenities mapped will be summed to tell us
how amenity dense a given area is from the point of a specific block.

*Note: only cultural amenities are considered, but future work should*
*expand to consider grocery stores, hospitals, employment and so forth as well.*


### Loading libraries
```{r include=TRUE, message=FALSE, warning=FALSE}

# Main
library(r5r)
library(sf)
library(data.table)

if (Sys.getenv("JAVA_HOME")!="") { Sys.setenv(JAVA_HOME="") }
library(rJava)

# Convenience
library(tidyverse)
library(glue)
library(rlist)

```


### 1. Setup the Network Graph

- Requires GTFS and OSM data to be present in the same directory.

```{r message=FALSE, warning=FALSE}

## Allocate 4G RAM to Java
options(java.parameters = "-Xmx6g")

## Build transport network, pointing to path where OSM and GTFS data are located
r5r_core <- setup_r5(data_path = '../../../data/1_raw/transit_and_osm_data/', verbose = FALSE)

```


### 2. Load origin/destination points

```{r, warning=FALSE}

origins <- fread(file.path("../../../data/2_clean/", "vancouver_db.csv"))
destinations <- fread(file.path("../../../data/2_clean", "vancouver_facilities.csv"))


# minor cleaning
origins <- origins[, c(1,3,4)]
colnames(origins)[1] <- 'id'
origins$id <- as.character(origins$id)  # numeric to char

# minor cleaning
destinations <- destinations[, c(1:3)]
destinations$lat <-  as.numeric(destinations$lat) # char to numeric
destinations$lon <-  as.numeric(destinations$lon) # char to numeric
destinations$id <- as.character(destinations$id)  # numeric to char
destinations <- destinations[complete.cases(destinations)] # remove NA rows

# Peek
head(origins)
head(destinations)

```


### 3. Compute All Amenities within 1km Walking Distance

```{r}

#departure_datetime <- as.POSIXct(glue("01-05-2021 00:00:00"), format="%d-%m-%Y %H:%M:%S")
walk_ttm <- travel_time_matrix(r5r_core = r5r_core,
                               origins = origins, destinations = destinations,
                              #departure_datetime = departure_datetime,
                              
                              # constraints
                              mode = c('WALK'),
                              max_walk_dist = 1000,
                              max_trip_duration = 120,
                              verbose = FALSE)

print('COMPLETED')
summary(walk_ttm)

```

*Count amenities accessible to every dissemination block*

```{r}


walk_ttm_agg <- walk_ttm %>%
                group_by(fromId) %>% 
                summarise(n_amenities = n())

summary(walk_ttm_agg)
head(walk_ttm_agg)

```

### Export
```{r}

write.csv(walk_ttm_agg, "../../../data/3_computed/amenity_density.csv",
          row.names = FALSE)

```

